Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-24 mr-review.Rmd
added placeholder for inSIDE assumption

@@

Note that $k$ indexes SNP instruments and ranges from 1 to $K$.

@@

@@

>>>>

InSIDE assumption

@@

Note that $k$ indexes SNP instruments and ranges from 1 to $K$ in Equation \@ref(eq:smm1).

@@

@bowden_consistent_2016, noting that consistency with IVW MR is not robust to 
invalid instrument use, developed a method for consistent MR analysis with  
weighted median-based estimators. Building on research from @han2008detecting, 
@bowden_consistent_2016 devised a weighted median-based estimator. Its breakdown point is 50\%, 
meaning that up to 50\% of SNP instruments can be invalid while maintaining consistency in estimation 
[@hampel1971general]. Specifically, let $\tilde \beta_j$ denote the $j^{th}$ ordered ratio estimate 
statistic, from least to greatest. If an odd number, say $2K +1$, of SNP instruments is used, then 
the simple median estimator is defined as the $K + 1$ ordered ratio estimate. For an even number, say 
$2K$, number of SNP instruments, choose the midpoint between the two middle-ranking ordered ratio 
estimates, $\frac{\tilde \beta_K + \tilde \beta_{K + 1}}{2}$.

Due to the inefficiency of the simple median estimator, @bowden_consistent_2016 define the weighted 
median estimator. To construct this estimator, order the ratio estimates from least to greatest, as 
with the simple median estimator. The weighted median estimator, then is the median of a distribution
defined by having $\tilde \beta_j$ as the $p_j = 100(\sum_{k = 1}^j - \frac{w_j}{2})$ percentile, 
where $w_j$ is the weight assigned to the $j^{th}$ ordered ratio estimate. In this setting, the 
simple median estimator is seen to be the weighted median estimator when all weights are equal.

@bowden_consistent_2016 also studied the penalized weighted median estimator. They defined the 
penalized weights, $w_j^* = w_j' \times \min(1, 20q_j)$, where $q_j$ is the p-value resulting from 
the comparison of $Q_j = w_j'(\tilde\beta_j - \hat\beta_{IVW})^2$ to 
a $\chi^2_1$ distribution [@bowden_consistent_2016;@cochran1950comparison]. The penalized 
weighted median estimator

weighted mode estimator

@@

### MR with allele scores

Another approach to integrating multiple SNP instruments into a MR framework is
through calculation of allele scores [@burgess_use_2013]. @burgess_use_2013 
calculate an unweighted score as the number of exposure-increasing alleles in 
the subject's genotypes. They also calculate a weighted allele score by using 
the exposure effects as weights. For example, a subject with $g_k$ copies of 
exposure-increasing alleles for SNP $k$ has an unweighted allele score $z = \sum_{k = 1}^K g_{ik}$ 
and a weighted allele score $z = \sum_{k = 1}^K w_kg_k$.

@burgess_use_2013 found that when weights are obtained from external data or from 
cross-validation or jackknife approaches applied to the analysis data, the allele score 
functions as a single instrumental variable and greatly diminishes the bias compared to that of the 
two-stage least squares estimator [@burgess_review_2017;@angrist1999jackknife].
